Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2018-03-28 LaTeX/S2Appendix_revised.tex
refining conclusions and reviewer 2 responses

%reference equations in appendices
\usepackage{xr}
\externaldocument{S1Appendix}
\externaldocument{main}
\externaldocument{S3Appendix}

@@

\noindent
A fifth equilibrium $(C)$ also exists where $A$ is present at an intermediate frequency on the Y background ($0<\hat{p}_{Y}^{\male}<1$). 
However, equilibrium $(C)$ is never locally stable when $r \approx 0$ and is therefore not considered further.
Thus, the Y background can either be fixed for the $a$ allele (equilibria ($A$) and ($B$)) or the $A$ allele (equilibria ($A'$) and ($B'$)).
The X background can then either be polymorphic (equilibria ($A$) and ($A'$)) or fixed for the alternative allele (equilibria ($B$) and ($B'$)).
Since equilibria $(A)$ and $(B)$ are equivalent to equilibria $(A')$ and $(B')$ with the labelling of $A$ and $a$ alleles interchanged, we discuss only equilibria $(A)$ and $(B)$, in which the Y background is fixed for the $a$ allele. 
If there is no haploid selection ($\alpha^{\Hermaphrodite}_\Delta=0$, $w_{A}^{\Hermaphrodite}=w_{a}^{\Hermaphrodite}=1$), these equilibria are equivalent to those found by~\cite{Lloyd1977} and~\cite{Otto2014}.

@@

Here, we assume that selection is weak relative to recombination ($s^\Hermaphrodite$, $t^\Hermaphrodite$, $\alpha_{\Delta}^\Hermaphrodite$ of order $\epsilon$). 
The maintenance of a polymorphism at the $\mathbf{A}$ locus then requires that

@@

\noindent
where $V_{A}=\bar{p}(1-\bar{p})$ is the variance in the frequency of $A$ and $D^\Hermaphrodite=\big{[} \bar{p}s^\Hermaphrodite+(1-\bar{p})h^\Hermaphrodite s^\Hermaphrodite\big{]} -\big{[} \bar{p}h^\Hermaphrodite s^\Hermaphrodite+(1-\bar{p}) \big{]}$ corresponds to the difference in fitness between $A$ and $a$ alleles in diploids of sex $\Hermaphrodite \in \{\female,\male\}$. 
The frequency of Y-bearing male gametes depends upon the difference in the frequency of the $A$ allele between X- and Y-bearing male gametes and the strength of meiotic drive in favour of the $A$ allele in males, $q=1/2+\alpha_{\Delta}^\male(\hat{p}^\male_Y-\hat{p}^\male_X)/2+O(\epsilon^3)$.
Without gametic competition or drive ($\alpha_{\Delta}^\Hermaphrodite=t^\Hermaphrodite=0$) our results reduce to those of~\cite{vanDoorn:2007eu}.

>>>>

%reference equations in appendices
\usepackage{xr}
\externaldocument{S1Appendix_revised}
\externaldocument{main_revised}
\externaldocument{S3Appendix_revised}

@@

\noindent
A fifth equilibrium $(C)$ also exists where $A$ is present at an intermediate frequency on the Y background ($0<\hat{p}_{Y}^{\male}<1$). 
However, equilibrium $(C)$ is never locally stable when $r \approx 0$ and is therefore not considered further.
Thus, the Y background can either be fixed for the $a$ allele (equilibria ($A$) and ($B$)) or the $A$ allele (equilibria ($A'$) and ($B'$)).
The X background can then either be polymorphic (equilibria ($A$) and ($A'$)) or fixed for the alternative allele (equilibria ($B$) and ($B'$)).
Since equilibria $(A)$ and $(B)$ are equivalent to equilibria $(A')$ and $(B')$ with the labelling of $A$ and $a$ alleles interchanged, we discuss only equilibria $(A)$ and $(B)$, in which the Y background is fixed for the $a$ allele. 
If there is no haploid selection ($\alpha^{\circ}_\Delta=0$, $w_{A}^{\circ}=w_{a}^{\circ}=1$), these equilibria are equivalent to those found by~\cite{Lloyd1977} and~\cite{Otto2014}.

@@

Here, we assume that selection is weak relative to recombination ($s^\circ$, $t^\circ$, $\alpha_{\Delta}^\circ$ of order $\epsilon$). 
The maintenance of a polymorphism at the $\mathbf{A}$ locus then requires that

@@

\noindent
where $V_{A}=\bar{p}(1-\bar{p})$ is the variance in the frequency of $A$ and $D^\circ=\big{[} \bar{p}s^\circ+(1-\bar{p})h^\circ s^\circ\big{]} -\big{[} \bar{p}h^\circ s^\circ+(1-\bar{p}) \big{]}$ corresponds to the difference in fitness between $A$ and $a$ alleles in diploids of sex $\circ \in \{\female,\male\}$. 
The frequency of Y-bearing male gametes depends upon the difference in the frequency of the $A$ allele between X- and Y-bearing male gametes and the strength of meiotic drive in favour of the $A$ allele in males, $q=1/2+\alpha_{\Delta}^\male(\hat{p}^\male_Y-\hat{p}^\male_X)/2+O(\epsilon^3)$.
Without gametic competition or drive ($\alpha_{\Delta}^\circ=t^\circ=0$) our results reduce to those of~\cite{vanDoorn:2007eu}.

### 2018-03-28 LaTeX/S3Appendix_revised.tex
refining conclusions and reviewer 2 responses

%reference equations in appendices
\usepackage{xr}
\externaldocument{S1Appendix}
\externaldocument{S2Appendix}
\externaldocument{main}

@@

Haploid selection generally expands the conditions under which neo-W alleles can spread within ancestral systems that have evolved tight linkage between the sex-determining locus and a selected locus ($r \approx 0$).
First, haploid selection can allow a polymorphism to be maintained when it would not under diploid selection alone (e.g., with directional selection in diploids). 
In cases of ploidally-antagonistic selection, where there is a balance between alleles favored in the haploid stage and the diploid stage, neo-W alleles - even if unlinked to the selected locus - can spread (\nameref{fig:regionPloidAntag}).
Second, even when diploid selection could itself maintain a polymorphism, haploid selection can increase the conditions under which transitions among sex-determining systems are possible.
Of particularly importance, when selection is sexually-antagonistic in diploids ($s^\female s^\male <0$ and $0<h^\Hermaphrodite<1$), an unlinked neo-W ($R=1/2$) cannot invade unless there is also haploid selection (see proof in \nameref{file:Mathematica}; Fig \ref{fig:SexAntagTighter} and \nameref{fig:SexAntagTighterMaleDrive}). 
More generally, haploid selection alters the conditions under which neo-W alleles can spread (compare \nameref{fig:regionMaleDrive}-\nameref{fig:regionFemaleGS} to Fig \ref{fig:regionplots}).

>>>>

%reference equations in appendices
\usepackage{xr}
\externaldocument{S1Appendix_revised}
\externaldocument{S2Appendix_revised}
\externaldocument{main_revised}

@@

Haploid selection generally expands the conditions under which neo-W alleles can spread within ancestral systems that have evolved tight linkage between the sex-determining locus and a selected locus ($r \approx 0$).
First, haploid selection can allow a polymorphism to be maintained when it would not under diploid selection alone (e.g., with directional selection in diploids). 
In cases of ploidally-antagonistic selection, where there is a balance between alleles favored in the haploid stage and the diploid stage, neo-W alleles - even if unlinked to the selected locus - can spread (\nameref{fig:regionPloidAntag}).
Second, even when diploid selection could itself maintain a polymorphism, haploid selection can increase the conditions under which transitions among sex-determining systems are possible.
Of particularly importance, when selection is sexually-antagonistic in diploids ($s^\female s^\male <0$ and $0<h^\circ<1$), an unlinked neo-W ($R=1/2$) cannot invade unless there is also haploid selection (see proof in \nameref{file:Mathematica}; Fig \ref{fig:SexAntagTighter} and \nameref{fig:SexAntagTighterMaleDrive}). 
More generally, haploid selection alters the conditions under which neo-W alleles can spread (compare \nameref{fig:regionMaleDrive}-\nameref{fig:regionFemaleGS} to Fig \ref{fig:regionplots}).
